---
title: "PA1_template"
author: "William Yip"
date: "Sunday, May 17, 2015"
output: html_document
---

Loading and preprocessing the data

```{r}
activity <- read.csv("./activity.csv")
summary(activity)
```

Histogram of total number of steps taken per day

```{r, echo=TRUE}
hist(tapply(activity$steps,activity$date,sum), main = paste("Histogram of Total Number of Steps Taken per Day"), xlab="Sum of Steps")
```

Calculate mean and median total number of steps taken per day

```{r, echo=TRUE}
meanSteps <- mean(tapply(activity$steps,activity$date,sum), na.rm=TRUE)
medianSteps <- median(tapply(activity$steps,activity$date,sum), na.rm=TRUE)
print(paste0("Mean total number of steps taken per day is: ", meanSteps), digits = 2)
print(paste0("Median total number of steps taken per day is: ", medianSteps))
```
Average daily activity pattern
```{r, echo=TRUE}
avgStepsInt <- tapply(activity$steps,activity$interval,mean, na.rm=TRUE)
plot(names(avgStepsInt), avgStepsInt, type="l", main = "Time Series Plot", xlab="5-minute Intervals", ylab="Avg Steps")

## which.max gives index number of the max 5-minute value, so to get the
## max value - which is a lable, make the label a numeric

print(paste0("Interval ", as.numeric(names(which.max(avgStepsInt))) , " contains the maximum number of steps."))
```
Imputing Missing Values
```{r, echo=TRUE}
print(paste0("There are ", sum(is.na(activity)) , " missing values in the dataset."))
```
Filling in missing values in the dataset

```{r, echo=TRUE}

activityNoNA <- activity
meanInterval <-tapply(activity$steps, activity$interval,mean, na.rm=TRUE)
for (i in which(is.na(activityNoNA)))
    {
    activityNoNA[i,1] <- meanInterval[((i-1)%%288)+1]
    }
```
Historgram with new Dataset
```{r, echo=TRUE}
hist(tapply(activityNoNA$steps,activityNoNA$date,sum), main = paste("Histogram of Total Number of Steps Taken per Day"), xlab="Sum of Steps")
```

differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}
weekend <- c("Sunday","Saturday")

dayWeek <- c()

## Check for day of week as a weekend, if not, it's a weekday
for (i in 1:length(activityNoNA$steps))
    {
    if (any(weekdays(as.Date(activityNoNA[i,2])) == weekend)) 
        {
            ##add "weekend"" to new column for that row
            dayWeek[i] <- "weekend"  }
            ##else add "weekday"" to new column for that row
        else { dayWeek[i] <- "weekday"
         }
    }
## Combine activity file with dayweek column

    activityNoNA <- cbind(activityNoNA,dayWeek)

```
Weekday/weekend charts
```{r, echo=TRUE}
library(lattice)

##split dataset into weekend rows and weekday rows
## change the resulting lists to dataframes

a <- split(activityNoNA, activityNoNA$dayWeek)
wkDays <- do.call(rbind.data.frame,a[1])
wkEnd <- do.call(rbind.data.frame,a[2])


avgStepsIntwkDay <- tapply(wkDays$steps,wkDays$interval,mean)
avgStepsIntwkEnd <- tapply(wkEnd$steps,wkEnd$interval,mean)


intPlotwd <- as.data.frame(cbind(avgStepsIntwkDay, as.numeric(as.character((names(avgStepsInt))))))
intPlotwe <- as.data.frame(cbind(avgStepsIntwkEnd, as.numeric(as.character((names(avgStepsInt))))))

df.intPlotwe <- cbind(intPlotwe, rep("weekend",288))
df.intPlotwd <- cbind(intPlotwd, rep("weekday",288))

colnames(df.intPlotwd) <- c("avgSteps", "interval","dayweek")
colnames(df.intPlotwe) <- c("avgSteps", "interval","dayweek")


intPlot <- rbind(df.intPlotwd, df.intPlotwe)


```